Patient-Specific Neural Field Modeling from Micro-Electrode Data

### 2010-09-22 ltx/data_paper/DataPaper.tex
changes to the kernel basis functions to show that we can switch the inner product and the convolution

\subsection{Parametrisation of Connectivity Kernel}
The connectivity kernel can be decomposed as 
\begin{equation}\label{eq:DefKernelDecomp}
	 w\left(\mathbf{r},\mathbf{r}'\right) \approx \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta},
\end{equation}
where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. By assuming a Gaussian isotropic connectivity structure, the kernel basis functions can be written as $\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$. We will assume that we know the parametric form of the connectivity basis functions from the correlation analysis, where the scaling parameters $\boldsymbol{\theta}$ are unknown.
Substituting equation~\ref{eq:DefKernelDecomp} into equation\ref{eq:ReducedForm} gives
\begin{equation}\label{eq:KernelDecompForm}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \boldsymbol{\phi}(\mathbf{r}) 
	 \int_\Omega \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta} f(\boldsymbol{\phi}^{\top} (\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} 
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
The connectivity kernel weights can be shifted outside the integral giving
\begin{equation}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \boldsymbol{\phi}(\mathbf{r}) 
	 \int_\Omega \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
Now taking the basis function vector inside the inner integrand gives
\begin{equation}\label{eq:ReucedFormWithKernelDecomp}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \int_\Omega \boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
This equation can be simplified by exploiting the isotropy of the kernel and, hence, the basis functions used to represent it where
\begin{equation}
	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi} (\mathbf{r}'-\mathbf{r}).
\end{equation}
To make the simplification, we define
\begin{equation}\label{eq:DefPsi}
	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
\end{equation}
which is a constant $L \times n_{\theta}$ matrix that is defined analytically, where $L$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions. Substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReucedFormWithKernelDecomp} gives
\begin{equation}
	\mathbf{x}_{t+1} = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t 
+ \mathbf{e}_t(\mathbf{r}).
\end{equation}
This substitution exploits the isotropy of the connectivity kernel basis functions, allowing us to swap the convolution in equation~\ref{eq:ReucedFormWithKernelDecomp}, which is computationally demanding, with the convolution in equation~\ref{eq:DefPsi}. This provides a dramatic increase in estimation speed, since the matrix $\boldsymbol\Psi(\mathbf{r}')$ can be calculated analytically. This is of great importance as it lowers the computational demands of the estimation algorithm.

>>>>

\subsection{Parametrisation of Connectivity Kernel}
The connectivity kernel can be decomposed as 
\begin{equation}\label{eq:DefKernelDecomp}
	 w\left(\mathbf{r},\mathbf{r}'\right) \approx \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta},
\end{equation}
where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. By assuming a Gaussian isotropic connectivity structure, the kernel basis functions can be written as $\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$. We will assume that we know the parametric form of the connectivity basis functions from the correlation analysis, where the scaling parameters $\boldsymbol{\theta}$ are unknown.
Substituting equation~\ref{eq:DefKernelDecomp} into equation\ref{eq:ReducedForm} gives
\begin{equation}\label{eq:KernelDecompForm}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \boldsymbol{\phi}(\mathbf{r}) 
	 \int_\Omega \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta} f(\boldsymbol{\phi}^{\top} (\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} 
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
The connectivity kernel weights can be shifted outside the integral giving
\begin{equation}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \boldsymbol{\phi}(\mathbf{r}) 
	 \int_\Omega \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
Now taking the basis function vector inside the inner integrand gives
\begin{equation}\label{eq:ReucedFormWithKernelDecomp}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \int_\Omega \boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
This equation can be simplified by exploiting the isotropy of the basis functions used to represent the kernel, and hence,
\begin{equation}
	\psi_i (\mathbf{r}-\mathbf{r}') = \psi_i (2c_i+\mathbf{r}'-\mathbf{r}).
\end{equation}
where $c_i$ is the center of the $i^{th}$ basis function of the kernel decomposition. To make the simplification, we define
\begin{equation}\label{eq:DefPsi}
	\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}  \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\psi_i (2c_i+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
\end{equation}
where $\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}$ denotes the $i^{th}$column of $\boldsymbol{\Psi}(\mathbf{r}')$ which is a constant $L \times n_{\theta}$ matrix that is defined analytically, where $L$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions. Substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReucedFormWithKernelDecomp} gives
\begin{equation}
	\mathbf{x}_{t+1} = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t 
+ \mathbf{e}_t(\mathbf{r}).
\end{equation}
This substitution exploits the isotropy of the connectivity kernel basis functions, allowing us to swap the convolution in equation~\ref{eq:ReucedFormWithKernelDecomp}, which is computationally demanding, with the convolution in equation~\ref{eq:DefPsi}. This provides a dramatic increase in estimation speed, since the matrix $\boldsymbol\Psi(\mathbf{r}')$ can be calculated analytically. This is of great importance as it lowers the computational demands of the estimation algorithm.
